---
title: "P8105 Homework 1"
author: "Maria Serafini"
output: github_document
---

## Problem 1

```{r setup, include=FALSE}
library(moderndive)
library(tidyverse)
```


__This chunk loads and describes the `early_january_weather` dataset from the moderndive library__ 

```{r early_january_weather data}
#loads dataset early_january_weather
data("early_january_weather")

# calculates and saves value for the mean, median, and standard deviation of temperature variable
mean_temp <- mean(pull(early_january_weather, temp))
median_temp <- median(pull(early_january_weather, temp))
std_temp <- sd(pull(early_january_weather, temp))

# calculates and saves value the mean, median, and standard deviation of temperature variable
mean_humid <- mean(pull(early_january_weather, humid))
median_humid <- median(pull(early_january_weather, humid))
std_humid <- sd(pull(early_january_weather, humid))
```

__The dataset `early_january_weather` includes meterological data for LGA, JFK, and EWR for the month of January 2013 and contains `r nrow(early_january_weather)` observations (rows) representing hourly measurements and `r ncol(early_january_weather)` variables (columns):__ 

* `origin` - weather station (LGA, JFK, EWR)
* `year` - year at time of recording (2013)
* `month` - month at time of recording (1)
* `day` - day of the month at time of recording (`r range(pull(early_january_weather, day))`)
* `hour` - hour of the day at time of recording (`r range(pull(early_january_weather, hour))`)
* `temp` - temperature (degrees F, _mean_: `r round(mean_temp, digits = 2)`, _median_: `r round(median_temp, digits = 2)`, _std_: `r round(std_temp, digits = 2)`)
* `dewp` - dewpoint (degrees F)
* `humid` - relative humidity (_mean_: `r round(mean_humid, digits = 2)`, _median_: `r round(median_humid, digits = 2)`, _std_: `r round(std_humid, digits = 2)`)
* `wind_dir` - wind direction (degrees)
* `wind_speed` - wind speed (mph) 
* `wind_gust` - wind gust speed (mph)
* `precip` - precipitation (inches)
* `pressure` - sea level pressure (milibars)
* `visib` - visibility (miles)
* `time_hour` - date and hour of recording as a POSIXct date

#
__This chunk makes a scatterplot of temperature by date/hour of recording__

```{r scatterplot}
# gives scatter plot of temperature by date/hour of recording and color coded based on relative humidity
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point()
```



