---
title: "Homework 1"
author: Maria Serafini
output: github_document
---

# Problem 1

```{r setup, include=FALSE}
library(moderndive)
library(tidyverse)
```

### This chunk loads the 'early_january_weather' dataset from the moderndive library

```{r early_january_weather data}
#loads dataset early_january_weather
data("early_january_weather")

# gives the size of the dataset by rows and columns
nrow(early_january_weather)
ncol(early_january_weather)

# calculates the mean temperature
mean_temp <- mean(pull(early_january_weather, temp))
```

### The dataset `early_january_weather` includes meterological data for LGA, JFK, and EWR for the month of January 2013.
* __Contains 358 observations representing hourly measurements and 15 variables:__ 
  * _origin_ - weather station (LGA, JFK, EWR)
  * _year_ - year at time of recording (2013)
  * _month_ - month at time of recording (1)
  * _day_ - day of the month at time of recording (1-15)
  * _hour_ - hour of the day at time of recording (0-24)
  * _temp_ - temperature (degrees F)
  * _dewp_ - dewpoint (degrees F)
  * _humid_ - relative humidity
  * _wind_dir_ - wind direction (degrees)
  * _wind_speed_ - speed (mph) 
  * _wind_gust_ - gust speed (mph)
  * _precip_ - precipitation (inches)
  * _pressure_ - sea level pressure (milibars)
  * _visib_ - visibility (miles)
  * _time_hour_ - date & hour of recording as a POSIXct date
* __The size of the dataset is `r nrow(early_january_weather)` rows by `r ncol(early_january_weather)` columns__
* __The mean temperature is `r round(mean_temp, digits = 2)`__