---
title: "Homework 1"
author: Maria Serafini
output: github_document
---

# Problem 1

```{r setup, include=FALSE}
library(moderndive)
library(tidyverse)
```

## This chunk loads the 'early_january_weather' dataset from the moderndive library

```{r early_january_weather data}
#loads dataset early_january_weather
data("early_january_weather")

# gives the size of the dataset 
nrow(early_january_weather)
ncol(early_january_weather)

# calculates the mean temperature
mean(pull(early_january_weather, temp))
```

## The dataset `early_january_weather` includes meterological data for LGA, JFK, and EWR for the month of January 2013.
* Contains 358 rows representing hourly measurements and 15 variables (origin, year, month, day, hour, temp, dewp, humid, wind_dir, wind_speed, wind_gust, precip, pressure, visib, time_hour) 
  * origin: weather station (LGA, JFK, EWR)
  * year, month, day, hour: time of recording
  * temp, dewp: temperature and dewpoint (degrees F)
  * humid: relative humidity
  * wind_dir, wind_speed, wind_gust: wind direction (degrees), speed and gust speed (mph)
  * precip: precipitation (inches)
  * pressure: sea level pressure (milibars)
  * visib: visibility (miles)
  * time_hour: date & hour of recording as a POSIXct date
* The size of the dataset is `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. 
* The mean temperature is `r mean(pull(early_january_weather, temp))`